Patients with Left Hemispheric Intracerebral Hemorrhage are Less Likely to Receive Neurosurgical Intervention

### 2024-09-27 manuscripts/manuscript.qmd
all models run again

Previous research involving patients with ischemic stroke provide evidence that those with right hemispheric injury present later to the hospital and may not be treated as frequently with thrombolytic therapy. [@dileggeImpactLesionSide2005; @foerchDifferenceRecognitionRight2005; @portegiesLeftsidedStrokesAre2015;  @liStrokeLateralizationLarge2021] It has also been noted that there is a hesitation among some practitioners to offer decompressive craniectomy in patients with left dominant-hemispheric ischemic infarcts3 despite literature demonstrating that long-term functional outcomes and health-related quality of life (HRQoL) are equivalent if not better for patients with left hemispheric injuries. [@deb-chatterjiSideMattersDifferences2022; @leggeNeglectingDifference2006; @beuscherFunctionalLongTermOutcome2017; @satoHigherMortalityPatients2015]

It is uncertain how the results of prior work in the ischemic stroke literature translate to patients experiencing acute ICH. We aimed to evaluate the effect of hemispheric lateralization of ICH on likelihood of receiving aggressive surgical intervention.

@@

@@

#### Prior Distribution Specification
Two sets of priors were applied for all models, both assuming no association of hemispheric laterality on the outcomes of interest: 1) default, weakly informative priors as defined by the 'brms' software package; 2) priors informed from review of previous literature. A description of the reasons for these priors is available in the supplementary material.

#### Computation and Description of the Posterior Distribution
We report the following metrics for the effect of hemispheric laterality on the primary and secondary outcomes: 1) median odds ratio (OR) of the posterior distribution with 95% credible interval; 2) probability of any difference (P[OR] > 1); 3) probability of a large difference (P[OR] > 1.2) as defined by the study authors; 4) percent of the posterior distribution outside of the region of practical equivalence as defined by the study authors (ROPE; 0.95 > P[OR] > 1.05). A description of the reasons for these thresholds is available in the supplementary material.

@@

```{r table1}
#| label: tbl-1
#| tbl-cap: "Baseline Characteristics"
#| echo: false

with_dir(here(), tar_read(table_1_aggressive))

@@

```{r table 2}
#| label: tbl-2
#| tbl-cap: "Association between hemispheric lateralization and aggressive care"
#| echo: false
#| warning: false

# gather odds ratios and 95% credible intervals for each model
summary <- modelsummary(models,
  exponentiate = TRUE,
  fmt = 2,
  estimate = c("aOR" = "estimate"),
  statistic = c("95% Credible Interval" = "{conf.low} - {conf.high}"),
  coef_omit = "^(?!b_ich_lat)",
  coef_rename = c("b_ich_lateralityRight" = "Right Hemisphere ICH"),
  gof_map = c("none"),
  shape = model ~ term + statistic,
)

aOR_CIs_combined <- as_tibble(summary@table_dataframe) |>
  rename("Outcome" = " ")

aOR_CIs_combined[1] <- c(
  "Neurosurgical Intervention",
  "External Ventricular Drain",
  "Days of Mechanical Ventilation",
  "Tracheostomy",
  "Withdrawal of Life-Sustaining Therapy",
  "Early Withdrawal of Life-Sustaining Therapy",
  "DNR Order"
)

# gather posterior probabilities of various magnitudes of difference for each outcome

neurosurgery <- models$"Neurosurgical Intervention" |>
  spread_draws(b_ich_lateralityRight) |>
  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
  summarize(
    or_1 = sum(ich_right_or > 1) / n(),
    or_1.1 = sum(ich_right_or > 1.1) / n(),
    or_1.2 = sum(ich_right_or > 1.2) / n(),
    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
  )

evd <- models$"External Ventricular Drain" |>
  spread_draws(b_ich_lateralityRight) |>
  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
  summarize(
    or_1 = sum(ich_right_or > 1) / n(),
    or_1.1 = sum(ich_right_or > 1.1) / n(),
    or_1.2 = sum(ich_right_or > 1.2) / n(),
    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
  )

vent <- models$"Days of Mechanical Ventilation" |>
  spread_draws(b_ich_lateralityRight) |>
  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
  summarize(
    or_1 = sum(ich_right_or > 1) / n(),
    or_1.1 = sum(ich_right_or > 1.1) / n(),
    or_1.2 = sum(ich_right_or > 1.2) / n(),
    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
  )

trach <- models$"Tracheostomy" |>
  spread_draws(b_ich_lateralityRight) |>
  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
  summarize(
    or_1 = sum(ich_right_or > 1) / n(),
    or_1.1 = sum(ich_right_or > 1.1) / n(),
    or_1.2 = sum(ich_right_or > 1.2) / n(),
    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
  )

wlst <- models$"Withdrawal of Life-Sustaining Therapy" |>
  spread_draws(b_ich_lateralityRight) |>
  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
  summarize(
    or_1 = sum(ich_right_or > 1) / n(),
    or_1.1 = sum(ich_right_or > 1.1) / n(),
    or_1.2 = sum(ich_right_or > 1.2) / n(),
    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
  )

early_wlst <- models$"Early Withdrawal of Life-Sustaining Therapy" |>
  spread_draws(b_ich_lateralityRight) |>
  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
  summarize(
    or_1 = sum(ich_right_or > 1) / n(),
    or_1.1 = sum(ich_right_or > 1.1) / n(),
    or_1.2 = sum(ich_right_or > 1.2) / n(),
    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
  )

dnr <- models$"DNR Order" |>
  spread_draws(b_ich_lateralityRight) |>
  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
  summarize(
    or_1 = sum(ich_right_or > 1) / n(),
    or_1.1 = sum(ich_right_or > 1.1) / n(),
    or_1.2 = sum(ich_right_or > 1.2) / n(),
    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
  )

# combine probability of OR > 1 (prob of difference)
prob_difference <- tibble(
  "Neurosurgical Intervention" = pluck(neurosurgery, 1),
  "External Ventricular Drain" = pluck(evd, 1),
  "Days of Mechanical Ventilation" = pluck(vent, 1),
  "Tracheostomy" = pluck(trach, 1),
  "Withdrawal of Life-Sustaining Therapy" = pluck(wlst, 1),
  "Early Withdrawal of Life-Sustaining Therapy" = pluck(early_wlst, 1),
  "DNR Order" = pluck(dnr, 1)
) |>
  pivot_longer(
    cols = everything(),
    names_to = "Outcome",
    values_to = "Probabilty of any difference (aOR > 1)"
  )

# combine % of posterior within the ROPE
rope <- tibble(
  "Neurosurgical Intervention" = pluck(neurosurgery, 4),
  "External Ventricular Drain" = pluck(evd, 4),
  "Days of Mechanical Ventilation" = pluck(vent, 4),
  "Tracheostomy" = pluck(trach, 4),
  "Withdrawal of Life-Sustaining Therapy" = pluck(wlst, 4),
  "Early Withdrawal of Life-Sustaining Therapy" = pluck(early_wlst, 4),
  "DNR Order" = pluck(dnr, 4)
) |>
  pivot_longer(
    cols = everything(),
    names_to = "Outcome",
    values_to = "% within ROPE"
  )

table_almost_2 <- left_join(aOR_CIs_combined, prob_difference, by = "Outcome") |>
  left_join(rope, by = "Outcome")

table_2 <- with_dir(here(), tar_read(table_2_aggressive))

tibble_2 <- table_2[[1]] |>
  select(c(label, stat_1, stat_2))

column_names <- c(
  "Neurosurgical Intervention",
  "External Ventricular Drain",
  "Days of Mechanical Ventilation",
  "Tracheostomy",
  "Withdrawal of Life-Sustaining Therapy",
  "Early Withdrawal of Life-Sustaining Therapy",
  "DNR Order"
)

tibble_2 <- tibble_2 |>
  rename(
    "Left Hemisphere" = "stat_1",
    "Right Hemisphere" = "stat_2",
    "Outcome" = "label"
  ) |>
  full_join(table_almost_2, by = "Outcome") |>
  gt()

tibble_2 |>
  fmt_number(columns = 6:7, decimals = 2) |>
  tab_footnote(
    footnote = "Adjusted for: age, admission GCS, ICH location, IVH, and study (as random intercept)",
    locations = cells_column_labels(columns = 4)
  ) |>
  tab_footnote(
    footnote = "Region of Practical Equivalence, defined as 0.95 > aOR > 1.05",
    locations = cells_column_labels(columns = 7)
  )

@@

predictions(models$"Neurosurgical Intervention", by = c("ich_laterality", "ich_location")) |>
  posterior_draws() |>
  ggplot(aes(x = draw, fill = ich_laterality)) +
  geom_density(alpha = .7) +
  facet_grid(~ich_location) +
  scale_fill_manual(values = c("#054224", "#ff8800")) +
  scale_y_continuous() +
  scale_x_continuous(
    limits = c(0, 0.2),
    breaks = seq(0, 0.20, 0.05),
    labels = scales::percent
  ) +
  labs(
    fill = "ICH Laterality",
    y = NULL
  ) +
  theme_ich()

>>>>

Previous research involving patients with ischemic stroke provide evidence that those with right hemispheric injury present later to the hospital and may not be treated as frequently with thrombolytic therapy. [@dileggeImpactLesionSide2005; @foerchDifferenceRecognitionRight2005; @portegiesLeftsidedStrokesAre2015;  @liStrokeLateralizationLarge2021] It has also been noted that there is a hesitation among some practitioners to offer decompressive craniectomy in patients with left dominant-hemispheric ischemic infarcts despite literature demonstrating that long-term functional outcomes and health-related quality of life (HRQoL) are equivalent if not better for patients with left hemispheric injuries. [@deb-chatterjiSideMattersDifferences2022; @leggeNeglectingDifference2006; @beuscherFunctionalLongTermOutcome2017; @satoHigherMortalityPatients2015]

Previous work has demonstrated mixed results with regard to hemispheric lateralization and outcomes in ICH, and there hasn't been much research examining if patients are treated differently based on the hemisphere affected. We aimed to evaluate the effect of hemispheric lateralization of ICH on likelihood of receiving aggressive surgical intervention as well as it's association with functional outcomes.

@@

Our primary functional outcome was modified Rankin score at 90 days, examined via ordinal regression.

@@

#### Prior Distribution Specification
Four sets of priors were applied for all models: both assuming no association of hemispheric laterality on the outcomes of interest: 1) default, weakly informative priors as defined by the 'brms' software package; 2) priors informed from review of previous literature. A description of the reasons for these priors is available in the supplementary material.

#### Computation and Description of the Posterior Distribution
We report the following metrics for the effect of hemispheric laterality on the primary and secondary outcomes: 1) median odds ratio (OR) of the posterior distribution with 95% credible interval; 2) probability of any difference (P[OR] > 1); 3) probability of a substantial difference (P[OR] > 1.2) as defined by the study authors; 4) percent of the posterior distribution outside of the region of practical equivalence as defined by the study authors (ROPE; 0.95 > P[OR] > 1.05). A description of the reasons for these thresholds is available in the supplementary material.

@@

@@

@@

predictions(models$"Neurosurgical Intervention", type = "response", by = c("ich_laterality")) |>
  posterior_draws() |>
  ggplot(aes(x = draw, fill = ich_laterality)) +
  geom_density(alpha = .7) +
  facet_grid(~ich_location) +
  scale_fill_manual(values = c("#054224", "#ff8800")) +
  scale_y_continuous() +
  scale_x_continuous(
    limits = c(0, 0.2),
    breaks = seq(0, 0.20, 0.05),
    labels = scales::percent
  ) +
  labs(
    fill = "ICH Laterality",
    y = NULL
  ) +
  theme_minimal()

predictions(models$"Neurosurgical Intervention", type = "response", by = c("ich_laterality", "ich_location")) |>
  posterior_draws() |>
  ggplot(aes(x = draw, fill = ich_laterality)) +
  geom_density(alpha = .7) +
  # facet_grid(~ich_location) +
  scale_fill_manual(values = c("#054224", "#ff8800")) +
  scale_y_continuous() +
  scale_x_continuous(
    limits = c(0, 0.2),
    breaks = seq(0, 0.20, 0.05),
    labels = scales::percent
  ) +
  labs(
    fill = "ICH Laterality",
    y = NULL
  ) +
  theme_minimal()
